Preparing your manuscript

### 2025-06-19 article/article.qmd
Something broke, now plots make sense again

@@

tos <- purrr::map(Sys.glob(here("data/tos/*hist*mean.nc")), function(f) {

  ReadNetCDF(f, vars = c(hist = var)) |> 
    _[, let(ssp585 = ReadNetCDF(file_scenario, vars = c(ssp585 = var), out = "vector")[[1]])] |> 
    _[, let(model = meta[[1]][["model"]],
            lat = round(lat, digits = 8),
            var = var,
            dt_dy_hist = metR:::.derv(hist, lat)*(180/pi/6371000),
            dt_dy_ssp585 = metR:::.derv(ssp585, lat)*(180/pi/6371000)), by = .(lon)] |> 
    _[, let(cos_lat = cos(lat*pi/180))] |> 
    _[, let(dt_dx_hist = metR:::.derv(hist, lon)*(180/pi/6371000)*cos_lat,
            dt_dx_ssp585 = metR:::.derv(ssp585, lon)*(180/pi/6371000)*cos_lat), by = .(lat)] |> 
    _[, let(change = sqrt(dt_dx_ssp585^2 + dt_dy_ssp585^2) / sqrt(dt_dx_hist^2 + dt_dy_hist^2),
            diff = sqrt(dt_dx_ssp585^2 + dt_dy_ssp585^2) - sqrt(dt_dx_hist^2 + dt_dy_hist^2))] |> 
    _[, .(time, lat, lon, hist, ssp585, model, var, change, diff)]

mask <- tos[var == "tos" & model == "NorESM2-MM", .(lon, lat, hist)] |> 
  _[, let(land = fifelse(!is.na(hist), "Land", "Sea"))] |> 
  _[, .(lon, lat, land)]

data <- rbindlist(list(delta, tas, tos)) |> 
  mask[x = _, on = .NATURAL]
  # _[, let(land = MaskLand(lon, lat))] |> 
  # _[, let(land = factor(land, labels = c("Sea", "Land")))]

@@

model_spread <- data |> 
  _[, let(serie = fcase(lat %between% c(48, 50), "50 - NH",
                        lat %between% c(38, 40), "40 - NH",
                        lat %between% c(28, 30), "30 - NH", 
                        lat %between% c(-50, -48), "50 - SH",
                        lat %between% c(-40, -38), "40 - SH",
                        lat %between% c(-30, -28), "30 - SH"))] |>
  _[!is.na(serie) & var == "deltat" & model %in% models_in_gadi] |>
  _[, .(mean_freq = weighted.mean(ssp585, cos(lat*pi/180))), by = .(serie, land, model)] |> 
  tidyr::separate(serie, into = c("lat", "hemisphere"), sep = " - ") |> 
  setDT() |> 
  _[, let(scenario = "SSP5-8.5")] |> 
  _[!(land == "Land" & hemisphere == "SH" & lat == 50)]

@@

dist |> 
  _[, let(serie = fcase(lat %between% c(48, 50) & land == "Sea", "50 - NH",
                        lat %between% c(38, 40) & land == "Sea", "40 - NH",
                        lat %between% c(28, 30) & land == "Sea", "30 - NH", 
                        lat %between% c(48, 50) & land == "Land", "50 - NH",
                        lat %between% c(38, 40) & land == "Land", "40 - NH",
                        lat %between% c(28, 30) & land == "Land", "30 - NH",
                        lat %between% c(-50, -48) & land == "Land", "50 - SH",
                        lat %between% c(-40, -38) & land == "Land", "40 - SH",
                        lat %between% c(-30, -28) & land == "Land", "30 - SH",
                        lat %between% c(-50, -48) & land == "Sea", "50 - SH",
                        lat %between% c(-40, -38) & land == "Sea", "40 - SH",
                        lat %between% c(-30, -28) & land == "Sea", "30 - SH"))] |>
  _[!is.na(serie)] |> 
  _[, .(mean_freq = weighted.mean(mean_freq, cos(lat*pi/180))), by = .(land, serie, percentile, scenario)] |> 
  dcast(land + serie + percentile ~ scenario, value.var = "mean_freq") |> 
  tidyr::separate(serie, into = c("lat", "hemisphere"), sep = " - ") |> 
  setDT() |> 
  ggplot(aes(factor(round(100-as.numeric(percentile), digits = 1)), ssp585 / historical)) +
  geom_col(aes(fill = lat), position = "dodge") +
  scale_fill_manual(values = c("#004C73", "#CC3311", "#ee8a33")) +
  scale_y_log10() +
  facet_grid(land~hemisphere) +
  labs(x = "Percentile", y = latex2exp::TeX("Fractional change of $\\Delta T$"),
       fill = "Latitude") +
  theme_minimal() +
  theme(legend.position = "inside",
        legend.position.inside = c(0.8, 0.93),
        legend.direction = "horizontal",
        legend.background = element_rect(fill = "#fbfbfb", color = "#fbfbfb"),
        panel.background = element_rect(fill = "#fbfbfb", color = NA),
        strip.text = element_text(size = 12))

>>>>

  # file_scenario <- Sys.glob(here(paste0("data/", meta[[1]][["var"]], "/*_", meta[[1]][["model"]], "_ssp585_mean.nc")))

@@

# tos <- purrr::map(Sys.glob(here("data/tos/*hist*mean.nc")), function(f) {
#   
#   message(f)
#   meta <- unglue::unglue(basename(f), patterns = c("{var}_{model}_historical_mean.nc"))
#   
#   file_scenario <- Sys.glob(here(paste0("data/", meta[[1]][["var"]], "/", 
#                                         meta[[1]][["var"]], "_", meta[[1]][["model"]], "_ssp585*")))
#   
#   var <- meta[[1]][["var"]]
#   
#   ReadNetCDF(f, vars = c(hist = var)) |> 
#     _[, let(ssp585 = ReadNetCDF(file_scenario, vars = c(ssp585 = var), out = "vector")[[1]])] |> 
#     _[, let(model = meta[[1]][["model"]],
#             lat = round(lat, digits = 8),
#             var = var,
#             dt_dy_hist = metR:::.derv(hist, lat)*(180/pi/6371000),
#             dt_dy_ssp585 = metR:::.derv(ssp585, lat)*(180/pi/6371000)), by = .(lon)] |> 
#     _[, let(cos_lat = cos(lat*pi/180))] |> 
#     _[, let(dt_dx_hist = metR:::.derv(hist, lon)*(180/pi/6371000)*cos_lat,
#             dt_dx_ssp585 = metR:::.derv(ssp585, lon)*(180/pi/6371000)*cos_lat), by = .(lat)] |> 
#     _[, let(change = sqrt(dt_dx_ssp585^2 + dt_dy_ssp585^2) / sqrt(dt_dx_hist^2 + dt_dy_hist^2),
#             diff = sqrt(dt_dx_ssp585^2 + dt_dy_ssp585^2) - sqrt(dt_dx_hist^2 + dt_dy_hist^2))] |> 
#     _[, .(time, lat, lon, hist, ssp585, model, var, change, diff)]
#   
# }) |> rbindlist() 
# 
# mask <- tos[var == "tos" & model == "NorESM2-MM", .(lon, lat, hist)] |> 
#   _[, let(land = fifelse(!is.na(hist), "Land", "Sea"))] |> 
#   _[, .(lon, lat, land)]

data <- rbindlist(list(delta, tas)) |> 
  # mask[x = _, on = .NATURAL]
  _[, let(land = MaskLand(lon, lat))] |>
  _[, let(land = factor(land, labels = c("Sea", "Land")))]

@@

model_spread <- data |> 
  _[var == "deltat" & model %in% models_in_gadi] |> 
  _[, let(serie = fcase(lat %between% c(48, 50), "50° - NH",
                        lat %between% c(38, 40), "40° - NH",
                        lat %between% c(28, 30), "30° - NH", 
                        lat %between% c(-50, -48), "50° - SH",
                        lat %between% c(-40, -38), "40° - SH",
                        lat %between% c(-30, -28), "30° - SH"))] |>
  _[!is.na(serie)] |>
  _[, .(mean_freq = weighted.mean(ssp585, cos(lat*pi/180))), by = .(serie, land, model)] |> 
  tidyr::separate(serie, into = c("lat", "hemisphere"), sep = " - ") |> 
  setDT() |> 
  _[, let(scenario = "SSP5-8.5")] |> 
  _[!(land == "Land" & hemisphere == "SH" & lat == "50°")]

@@

dist |> 
  _[, let(serie = fcase(lat %between% c(48, 50) & land == "Sea", "50° - NH",
                        lat %between% c(38, 40) & land == "Sea", "40° - NH",
                        lat %between% c(28, 30) & land == "Sea", "30° - NH", 
                        lat %between% c(48, 50) & land == "Land", "50° - NH",
                        lat %between% c(38, 40) & land == "Land", "40° - NH",
                        lat %between% c(28, 30) & land == "Land", "30° - NH",
                        lat %between% c(-50, -48) & land == "Land", "50° - SH",
                        lat %between% c(-40, -38) & land == "Land", "40° - SH",
                        lat %between% c(-30, -28) & land == "Land", "30° - SH",
                        lat %between% c(-50, -48) & land == "Sea", "50° - SH",
                        lat %between% c(-40, -38) & land == "Sea", "40° - SH",
                        lat %between% c(-30, -28) & land == "Sea", "30° - SH"))] |>
  _[!is.na(serie)] |> 
  _[, .(mean_freq = weighted.mean(mean_freq, cos(lat*pi/180))), by = .(land, serie, percentile, scenario)] |> 
  dcast(land + serie + percentile ~ scenario, value.var = "mean_freq") |> 
  _[, let(historical = fifelse(serie  == "50° - SH" & land == "Land", 0, historical),
          ssp585 = fifelse(serie  == "50° - SH" & land == "Land", 0, ssp585))] |> 
  tidyr::separate(serie, into = c("lat", "hemisphere"), sep = " - ") |> 
  setDT() |> 
  ggplot(aes(factor(round(100-as.numeric(percentile), digits = 1)), ssp585 / historical)) +
  geom_col(aes(fill = lat), position = "dodge") +
  scale_fill_manual(values = c("#004C73", "#CC3311", "#ee8a33")) +
  scale_y_log10() +
  facet_grid(land~hemisphere) +
  labs(x = "Percentile", y = latex2exp::TeX("Fractional change of $\\Delta T$"),
       fill = "Latitude") +
  theme_minimal() +
  theme(legend.position = "inside",
        legend.position.inside = c(0.8, 0.93),
        legend.direction = "horizontal",
        legend.background = element_rect(fill = "#fbfbfb", color = "#fbfbfb"),
        panel.background = element_rect(fill = "#fbfbfb", color = NA),
        strip.text = element_text(size = 12))
